Transferability of clinical prediction models for early trauma care

### 2019-07-02 ManuscriptMarkdown.Rmd
Comment on manuscript

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in 
substantial uncertainty, with some variation between different data sets. Comparing 
transferred performance with the reciveing sample validation performance resulted in 
further uncertainty.

The most notable effects on model performance following model
transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
model, with a mistriage rate increase by `r substr(format(highest.comp.mistriage), 1, 5)`. 
Mainly contributing to this was a marked increase in overtriage. 
For this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
a high level of uncertainty and that our findings are compatible with a `r UI.top.bot[1]` 
decrease to a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]` 
would signifiy an improvement in model performance. The higher
95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.

The transfer of clinical prediction models in trauma care has not previously
been studied extensively . In 2016, Gerdin et al. found model transfer to
adversely affect model performance in terms of calibration, but that this could
be improved by updating the model [@RN27]. This study did not examine model
performance in terms of calibration, neither were the effects of updating the
model explored.

@@

## Strengths and limitations
This study design realistically reflects potential transfers: clinical
prediction models for trauma are usually developed in large metropolitan centres
and then transferred as clinical recommendations to minor non-metropolitan
centres.

Originally all single centres with a valid number of events were to constitute
a data set, but only one centre provided a sufficient number of events. Leading to 
the loss of the data set: Individual centres. This data set might have provided
results that could reflect realistic situations as major trauma centres regularly 
exchange data.

@@

## Clinical applications
This study shows that mistriage can potentially increase following
model transfer, and more prominently in certain cases. This supports 
the view that model transfers should be done with caution, as 
performance might decline.

Recognizing these risks and updating models accordingly will lead to increased
accuracy during trauma patient triage. Potentially saving lives by minimizing
undertriage, and likely save resources by minimizing overtriage. With registries
such as SweTrau, analysis of model transfer could be performed continuously to
optimize models currently in use and advise which model transfers might be
detrimental (or potentially beneficial).

## Future studies
Studies on model transfer can be performed in different settings, e.g. assessing
the transferability between different but similar countries, such as Sweden and
Norway. Also, increasing the number of patients would increase the power of the
results, one way of doing this would be to replicate this study with the latest
data in the SweTrau, as many hospitals have joined recently leading to an
increased number of patients registered.

@@

## Conclusion
Model transfer resulted in large and varying levels of uncertainty. Model transfer 
can therefore be unpredictable, and potentially lead to increased mistriage, resulting 
in increased patient mortality and a misappropriation of resources, due to increased 
undertriage and overtriage, respectively. Transfer of the `r colnames(highest.comp.mistriage)` 
model lead to an increased mistriage rate of `r highest.comp.mistriage`. This 
observed effect was also associated with considerable uncertainty, meaning mistriage might slightly
decrease, but would more likely increase, potentially substantially. Recognizing 
the limitations of this study, the authors believe further studies are warranted due
to the potential economic costs and patient consequences of poor model
transfers.

>>>>

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in
substantial uncertainty in its effect on model performance. Comparing
transferred performance with the reciveing sample validation performance
resulted in **further uncertainty** _(Not very clear what you mean here, please
try to clarify.)_.

The most notable effects on model performance following model transfer were seen
after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
resulted in an increase of the mistriage rate by 
`r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
a marked increase in overtriage.  For this transfer the 95% UI ranged from 
`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty and
that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
`r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
would signifiy an improvement in model performance. The higher 95% UI of `r
UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.

The transfer of clinical prediction models in trauma care has not previously
been studied extensively. In 2016, Gerdin et al. found model transfer to
adversely affect model performance in terms of calibration, but that this could
be improved by updating the model [@RN27]. This study did not examine model
performance in terms of calibration, neither were the effects of updating the
model explored.

@@

## Strengths and limitations
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres and then transferred as clinical
recommendations to minor non-metropolitan centres. The results should therefore
be of practical importance.

**One limitation is that** _(Try to highlight whether you disucss a strength or
limitation)_ originally all single centres with a valid number of events were to
constitute a data set, but only one centre provided a sufficient number of
events. Leading to the loss of the data set: Individual centres. This data set
might have provided results that could reflect realistic situations as major
trauma centres regularly exchange data.

@@

## Clinical applications
This study shows that mistriage can potentially increase following model
transfer, and more prominently in certain cases. This supports the view that
model transfers should be done with caution, as performance might
deteriorate. Recognizing this risks and updating models accordingly should lead
to increased accuracy during trauma patient triage. Potentially saving lives by
minimizing undertriage, and likely save resources by minimizing overtriage. With
registries such as SweTrau, analysis of model transfer could be performed
continuously to optimize models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

## **Future studies** _(Is this heading required by the journal? If not I suggest we remove these two paragraphs)_
Studies on model transfer can be performed in different settings, e.g. assessing
the transferability between different but similar countries. Also, increasing
the number of patients would increase the power of the results, one way of doing
this would be to replicate this study with the latest data in the SweTrau, as
many hospitals have joined recently leading to an increased number of patients
registered.

@@

## Conclusion
Model transfer resulted in large and varying levels of uncertainty in its effect
on mistriage. Model transfer can therefore be unpredictable, and potentially
lead to increased mistriage, **resulting in increased patient mortality and a
misappropriation of resources, due to increased undertriage and overtriage,
respectively** _(I suggest you move this part to clinical applications and that
you try to express these potential effects a bit more carefully)_. Transfer of
the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage
rate of `r highest.comp.mistriage`. This observed effect was also associated
with considerable uncertainty, meaning mistriage might slightly decrease, but
would more likely increase, potentially substantially. Recognizing the
limitations of this study, the authors believe further studies are warranted due
to the potential economic costs and patient consequences of poor model
transfers.
